---
title: "EDS 222: Homework 4"
format: 
  html:
    toc: true
---

```{r}
#| label: setup
#| include: false

# Load required packages
library(tidyverse)

```

## Human-bear conflicts

For this homework you are studying the probability of human-bear conflicts in California campgrounds. The data on workbench for this homework assignment includes a CSV file with 1,243 (fictional) records of California campgrounds describing whether or not they reported a human-bear conflict and some variables that could be influencing the probability of a conflict.

```{r}
#| label: load-data



```

## Seasonal differences?

An employee of California State Parks wants to know if human-bear conflicts are more common in the summer than the fall. Use a randomization test to answer their question.

### Questions

1.  What are the null and alternative hypotheses?
2.  What's the appropriate test statistic?
3.  Write the code for a randomization test. Remember to subset the data to the two seasons of interest!
4.  What's your p-value?
5.  How would you respond to your colleague at California State Parks?

## Probability of conflict

Your colleague shared your feedback with their supervisor and now they want more details. Specifically, they want to know how human disturbance and bear habitat influence the probability of a human-bear conflict. In the available data, you have variables for `road_density` (in km road / km^2^ area) and `dist_to_forest` (in km), which quantify human disturbance and bear habitat, respectively.

### Questions

1.  What model formulation would you choose to answer this question?
2.  Fit a logistic regression model to the available data.
3.  Examine the p-values for the "road density" and "distance to forest" coefficients. Are they significant at $\alpha=0.05$?
4.  Visualize your model results. Plot distance to forest on the x-axis and predicted probability of human-bear conflict on the y-axis. For your predicted probabilities, assume a road density of 8 km / km^2^. Facet your plot by season.
5.  California State Parks decides to only allocate human-bear conflict prevention resources (e.g., visitor education and ranger time) to campgrounds with a 20% probability of human-bear conflict in the summer. How close to a forest would a campground have to be to receive resources (assuming road density of 8 km / km^2^)?

## Omitted variable bias

Now that your interest in bear ecology is piqued, you start reading more of the literature. You find a paper with convincing evidence that bears are drawn to human food waste, and become concerned that this omitted variable may be biasing your conclusions from the previous model.

### Questions

1.  Your dataset includes `food_subsidy`, which is the amount of food found at the campground in non-bear-safe receptacles (in kg). Does human food waste correlate with road density and/or distance to forest? What does that mean for the risk of omitted variable bias?

2.  Add `food_subsidy` to your model. Is its coefficient significant? What happened to the estimates, standard errors, and p-values for the road density and distance to forest coefficients?

3.  Update your model visualization from the last section. Add prediction lines representing three different food subsidy levels: the 1Q, median, and 3Q values. Keep the road density at 8 km / km^2^. Qualitatively, how do these predictions differ from the earlier model's?

4.  What's the 95% confidence interval for the food subsidy coefficient? How would you interpret this CI?

5.  You share your updated model with your California State Parks colleague. They respond by asking, "I'm confused, didn't you say wildlife conflicts were caused by road density and forest proximity? What changed?" How would you respond?

## Providing recommendations

The managers of three campgrounds hear about your work and ask for to take action to reduce food subsidies. These campgrounds have the following characteristics:

| Campground | Distance to forest (km) | Road density (km / km^2^) | Food subsidy (kg) |
|----|----|----|----|
| A | 3 | 12 | 1.3 |
| B | 4 | 4 | 0.5 |
| C | 5 | 7 | 2.1 |

### Questions

1.  The managers ask, what are the predicted probabilities of human-bear conflicts in summer at their campgrounds?

2.  Do these probabilities exceed the California State Parks threshold of 20%? If so, how much would they need to reduce their food subsidies by to get below the threshold?

3.  How different would your predictions be if you had used the model without food subsidies?
